Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-11-04 Manuscript/manuscript.tex
Small correction to the Gamma values in MS.

\begin{table}[tb!]
  \caption{Relative surface excess of calcium with respect to water, 
  and the ratios of the amount of calcium cations 
  bound to phosphate or carbonyl moieties 
  to the amount of such cations bound to both moieties in a POPC bilayer 
  from various simulation models. 
  The CaCl$_2$ concentration of the ions in water before adsorption to the phospholipid bilayer
  is 350mM in the analyzed systems. \label{tab:binding}}
  \todo{Add columns with concentrations of calcium -- bulk/molar fraction.}
  \todo{Changing symbol $P$ to $R$ or $r$ might reflect better that it is rather a ratio, and not that much a probability.}
  \begin{tabular}{l|c c c}
    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $P^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $P^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
    \hline
    ECC-lipids            &  $0.06 \pm 0.01 $            &  99\%   &    25\%    \\
    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
    Lipid14/Dang          &  $ 0.3 \pm 0.03 $            &  -        &    -       \\
    Lipid14/ECC-ions      &  $ \simeq 0.4   $            &  -        &    -       \\
  \end{tabular}
\end{table}

Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation 
models were quantified by calculating the relative surface excess of calcium to water molecules, $\Gamma_{\rm Ca}^w$, from Eq.~\ref{surfexcess}.
%using the bulk concentrations determined from the density profiles in Fig.~\ref{fig:cacl-dens}. 
The values of $\Gamma_{\rm Ca}^w$, shown in Table~\ref{tab:binding},
were calculated from simulations with the same $350\,\mathrm{mM}$ concentration of CaCl$_2$ with 
respect to water, i.e., the concentration before solvating the lipids in experiments.
As expected from the changes of the lipid headgroup order parameters in Fig. \ref{fig:delta_ordPar_CaCl}, 
the relative surface excess for the ECC-lipids model, $\Gamma_{\rm Ca}^w = (0.06 \pm 0.01) \rm{nm}^{-2}$,
is significantly smaller than for the Lipid14 model with \AA{}qvist ions, $\Gamma_{\rm Ca}^w = (0.13 \pm 0.01) \rm{nm}^{-2}$,
or with Dang ions, $\Gamma_{\rm Ca}^w = (0.30 \pm 0.03) \rm{nm}^{-2}$,
or with ECC-ions,  $\Gamma_{\rm Ca}^w \simeq 0.4 \rm{nm}^{-2}$,
where the large uncertainity arises from a too small simulation box (Fig.~\ref{fig:cacl-dens}).
Interestingly, the relative surface excess of 
\ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) 
%calculated using density profiles in Fig.~\ref{fig:cacl-dens}
is qualitatively different from \ce{CaCl2}
having a negative value $\Gamma_i^w = (-0.11 \pm 0.01) \rm{nm}^{-2}$ 
meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface. 
This is in contrast to the most of the available lipid force fields,
which predict a specific binding of sodium to
PC lipid bilayers \cite{catte16}.
%\todo{JOE: Add references to NaCl simulation papers that show peaks.
%Add a simple note that there is NBFIX trick in Charmm36 (?).} 
%SAMULI: I think that it is enough to cite NMRlipids II here, which shows all the results
%
%This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
%which show the density profile of ions and water along the membrane normal. 
%
%The density profiles show
%a larger Ca$^{2+}$ density peak in lipid headgroup region for
%the Lipid14 model with Dang and ECC-ions than for the ECC-lipids model.

>>>>

\begin{table}[tb!]
  \caption{Relative surface excess of calcium with respect to water, 
  and the ratios of the amount of calcium cations 
  bound to phosphate or carbonyl moieties 
  to the amount of such cations bound to both moieties in a POPC bilayer 
  from various simulation models. 
  The CaCl$_2$ concentration of the ions in water before adsorption to the phospholipid bilayer
  is 350mM in the analyzed systems. \label{tab:binding}}
  \todo{Add columns with concentrations of calcium -- bulk/molar fraction.}
  \todo{Changing symbol $P$ to $R$ or $r$ might reflect better that it is rather a ratio, and not that much a probability.}
  \begin{tabular}{l|c c c}
    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $P^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $P^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
    \hline
    ECC-lipids            &  $0.06 \pm 0.01 $            &  99\%   &    25\%    \\
    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
    Lipid14/Dang          &  $0.23 \pm 0.03 $            &  -        &    -       \\
    Lipid14/ECC-ions      &  $0.35 \pm 0.11 $            &  -        &    -       \\
  \end{tabular}
\end{table}

Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation 
models were quantified by calculating the relative surface excess of calcium to water molecules, $\Gamma_{\rm Ca}^w$, from Eq.~\ref{surfexcess}.
%using the bulk concentrations determined from the density profiles in Fig.~\ref{fig:cacl-dens}. 
The values of $\Gamma_{\rm Ca}^w$, shown in Table~\ref{tab:binding},
were calculated from simulations with the same $350\,\mathrm{mM}$ concentration of CaCl$_2$ with 
respect to water, i.e., the concentration before solvating the lipids in experiments.
As expected from the changes of the lipid headgroup order parameters in Fig. \ref{fig:delta_ordPar_CaCl}, 
the relative surface excess for the ECC-lipids model, $\Gamma_{\rm Ca}^w = (0.06 \pm 0.01) \rm{nm}^{-2}$,
is significantly smaller than for the Lipid14 model with \AA{}qvist ions, $\Gamma_{\rm Ca}^w = (0.13 \pm 0.01) \rm{nm}^{-2}$,
or with Dang ions, $\Gamma_{\rm Ca}^w = (0.23 \pm 0.03) \rm{nm}^{-2}$,
or with ECC-ions,  $\Gamma_{\rm Ca}^w (0.35 \pm 0.11) \rm{nm}^{-2}$,
where the large uncertainity arises from a too small simulation box (Fig.~\ref{fig:cacl-dens}).
Interestingly, the relative surface excess of 
\ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) 
%calculated using density profiles in Fig.~\ref{fig:cacl-dens}
is qualitatively different from \ce{CaCl2}
having a negative value $\Gamma_i^w = (-0.11 \pm 0.01) \rm{nm}^{-2}$ 
meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface. 
This is in contrast to the most of the available lipid force fields,
which predict a specific binding of sodium to
PC lipid bilayers \cite{catte16}.
%\todo{JOE: Add references to NaCl simulation papers that show peaks.
%Add a simple note that there is NBFIX trick in Charmm36 (?).} 
%SAMULI: I think that it is enough to cite NMRlipids II here, which shows all the results
%
%This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
%which show the density profile of ions and water along the membrane normal. 
%
%The density profiles show
%a larger Ca$^{2+}$ density peak in lipid headgroup region for
%the Lipid14 model with Dang and ECC-ions than for the ECC-lipids model.
